---
title: "BIOL343 Coding Challenge 1"
author: "Group # — names and student numbers in the chunk below"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# BIOL343 Coding Challenge 1


## Group Members

```{r group_members}
# Name - Queen's Student Number

# 1.Ella Sha - 20457793
# 2.

```

# Coding Log — start (every work session)

Run these two lines in the **Console** (not in this chunk) each time you sit down to work on this challenge. The chunk is set to `eval=FALSE`, so it is skipped when you knit.

```{r positpal_start, eval=FALSE}
library(positpal)
positpal::start("BIOL343-CC1-template.Rmd")
```

# Part 1 — Import and Inspect the Dataset

## 1. Import the Data

The dataset is located in the `data` folder and is named:

`A1_Data_Excel.csv`

Use `read.csv()` to import the file and save it as an object called `assn1_data`.

```{r import_data}

# assn1_data <- read.csv("")   # fill in the path, then delete the leading #
assn1_data<-read.csv("~/Desktop/343 Data analysis/A1_Data_Excel.csv",header=T)
```

---

## 2. Inspect the Data

Use the following functions to examine the dataset:

- `class()`
- `dim()`
- `str()`
- `head()`
- `tail()`

```{r inspect_data}

# class()
class(assn1_data)

dim(assn1_data)


str(assn1_data)


head(assn1_data)


tail(assn1_data)

```

---

## 3. Clean the Dataset

Based on your inspection, retain the appropriate rows and columns.

Save the cleaned dataset as `assn1_data_mod`.

```{r clean_data}

assn1_data_mod <- assn1_data[1:46,1:8]

View(assn1_data_mod)
```

---

## 4. Calculate Ovary Area

Create a new variable called `ovary_area` using:

\[
\text{ovary area} = \text{ovary width} \times \text{ovary height}
\]

```{r ovary_area}

assn1_data_mod$ovary_area <- assn1_data_mod$Ovary.Width..mm. * assn1_data_mod$Ovary.Height..mm.
View(assn1_data_mod)
```

Use `head()` to confirm that the new column was created.

```{r check_ovary_area}

# Check the first few rows.
head(assn1_data_mod)
```

# Part 2 — Simulate and Combine Data

You will now create a small simulated dataset with the same variables as the original dataset.

For continuous variables such as ovary width and ovary height, generate values from a **normal distribution** using `rnorm()` and flower number using `rpois()`. Use a mean and standard deviation similar to those in the previously loaded `assn1_data_mod` dataset. You can calculate these values from the original data using `mean()` and `sd()`.

Before generating any random values, use `set.seed()` so that your simulated results can be reproduced.

Use the seed value `343`.

```{r set-seed}

set.seed(343)


```

> Run `set.seed(343)` before generating your simulated values. Using the same seed will produce the same random values each time the code is run.

## 5. Create Vectors

Create vectors for **10 new observations**.

### Population Type

Create a vector called `sim_population_type` containing `"A"` and `"S"` values.

```{r simulate-population}
set.seed(343)
 sim_population_type <- sample(c("A","S"), 10, replace=TRUE)
sim_population_type

```

### Ovary Width

Create a numeric vector called `sim_ovary_width` containing 10 simulated ovary-width values.

Use `rnorm()` and choose the mean and standard deviation based on the ovary-width values in `assn1_data_mod`.

```{r simulate-width}
set.seed(343)

#mean of ovary_width from assn1_data_mod
mean(assn1_data_mod$Ovary.Width..mm.)

#sd of ovary_width from assn1_data_mod
sd(assn1_data_mod$Ovary.Width..mm.)

 sim_ovary_width <- rnorm(10,3.424565,0.2924054)
sim_ovary_width
```

### Ovary Height

Create a numeric vector called `sim_ovary_height` containing 10 simulated ovary-height values.

Use `rnorm()` and choose the mean and standard deviation based on the ovary-height values in `assn1_data_mod`.

```{r simulate-height}
set.seed(343)
#mean of ovary_height from assn1_data_mod
mean(assn1_data_mod$Ovary.Height..mm.)

#sd of ovary_height from assn1_data_mod
sd(assn1_data_mod$Ovary.Height..mm.)

 sim_ovary_height <- rnorm(10,3.025,0.3584395)
sim_ovary_height
```

### Flower Number

Create a numeric vector called `sim_flower_number` containing 10 simulated flower-count values.

```{r simulate-flowers}

set.seed(343)
#mean of flower_number from assn1_data_mod
mean_F=mean(assn1_data_mod$Flower.Number,na.rm = TRUE)

#sd of flower_number from assn1_data_mod
sd_F=sd(assn1_data_mod$Flower.Number,na.rm = TRUE)


sim_flower_number <- c(rpois(10,lambda = mean_F))
sim_flower_number
```

---

## 6. Create a Simulated Data Frame

Combine your four vectors into a data frame called `sim_data`.

Make sure the column names match the corresponding columns in `assn1_data_mod`.

```{r create-sim-data}
sim_ovary_area <- sim_ovary_width * sim_ovary_height
sim_data  <- data.frame( Ovary.Width..mm. = sim_ovary_width,Ovary.Height..mm. = sim_ovary_height, Flower.Number= sim_flower_number, Population.Type=sim_population_type,ovary_area=sim_ovary_area)
sim_data
```

Calculate `ovary_area` for the simulated observations.

```{r simulated-ovary-area}

sim_data$ovary_area <- sim_data$Ovary.Width..mm. * sim_data$Ovary.Height..mm.

```

Use `str()` and `head()` to inspect your simulated dataset.

```{r inspect-sim-data}

str(sim_data)


head(sim_data)

```

---

## 7. Combine the Datasets

Combine the original cleaned dataset and the simulated dataset using `rbind()`.

Save the combined dataset as `combined_data`.

```{r combine-data}

combined_data <- rbind(assn1_data_mod[c("Population.Type","Ovary.Width..mm.","Ovary.Height..mm.","Flower.Number","ovary_area")],sim_data)
```

Check the dimensions of the combined dataset.

```{r combined-dimensions}

# Check the dimensions.
dim(combined_data)
```

# Part 3 — Basic Summary Statistics

## 8. Flower Number Summary Statistics

Calculate the following for `flower_number` in `combined_data`:

- Minimum
- Maximum
- Mean
- Median
- Standard deviation

```{r flower-summary}

# Minimum
min(combined_data$Flower.Number, na.rm = TRUE)

# Maximum
max(combined_data$Flower.Number, na.rm = TRUE)

# Mean
mean(combined_data$Flower.Number, na.rm = TRUE)

# Median
median(combined_data$Flower.Number, na.rm = TRUE)

# Standard deviation
sd(combined_data$Flower.Number, na.rm = TRUE)
```

---

## 9. Use `summary()`

Use `summary()` on both `ovary_area` and `flower_number`.

```{r summary-function}

# Ovary area
summary(combined_data$ovary_area)

# Flower number
summary(combined_data$Flower.Number)
```

**Question 1:** Which statistic that you calculated manually is not included in the output from `summary()`?

**Answer:**  
standard deviation is not included in the summary.
---

## 10. Compare Population Types

Create two subsets from `combined_data`:

- `asexual_data` for population type `"A"`
- `sexual_data` for population type `"S"`

```{r population-subsets}

asexual_data <- subset(combined_data,Population.Type=="A")
asexual_data

sexual_data <- subset(combined_data,Population.Type=="S")

```

Calculate the mean ovary area for each group.

```{r compare-populations}

# Mean ovary area for A
mean(asexual_data$ovary_area, na.rm=TRUE)

# Mean ovary area for S
mean(sexual_data$ovary_area, na.rm=TRUE)
```

**Question 2:** Which population type has the larger mean ovary area?

**Answer:**  The Asexual population has a larger mena ovary area.

---

## 11. Visualization of ovary area

Use `ggplot2` to create a figure of your choice comparing ovary area between population types (for example a boxplot, or histograms with `facet_wrap()`).

Your figure should include:

- `theme_classic()`
- appropriate axis labels (with units)
- a descriptive title
- a legend if necessary

```{r load_the_library}
install.packages("ggplot2")
library(ggplot2)


```

```{r visualization_of_ovary_area}
ggplot(aes(x = Population.Type, y = ovary_area),data=combined_data) + geom_boxplot(aes(fill=Population.Type)) + xlab("Population Type")+ ylab("Ovary Area(mm^2)") + labs(title = "Boxplot of ovary area for sexual and asexual population")

```

# Coding Log — export and upload (when you finish)

When the group's document is finished, **each group member** runs these lines in the Console, then uploads the `.zip` it creates to the **"CC1 — Coding Log (individual)"** dropbox on OnQ.

```{r positpal_end, eval=FALSE}
positpal::stop_capture()
positpal::export_capture()
# creates my-capture-YYYYMMDD-HHMMSS.zip in this folder
```
